### 2017-10-30 manuscript.rmd
fixed - plots are now centered

In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids. Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix and phylogeny due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes). ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\gamma$ model of amino acid substitution, and bootstrap values were estimated using 1000 replicates. BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112]. Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model. The CAT-GTR and CAT-poisson models did not converge after a long CPU time, and only the results from the CAT-poisson model are included here.

@@

The Tree:
```{r ultratree, echo=F, message=F, warning=F}
par(ask=F)
par(mfrow=c(1,1))
plotTree(ultratree, lwd = 4)

Stochastic trait mapping:
```{r simmap_making, echo=F, message=F, warning=F}
#Make simmap objects
print("SIMMAP Sex Distribution")
make.simmap(ultratree, sex, nsim = 100, message=FALSE) -> sex_sim
print("SIMMAP Palpons")
make.simmap(ultratree, palpons, nsim = 100, message=FALSE) -> palpons_sim
print("SIMMAP Tentilla")
make.simmap(ultratree, tentilla, nsim = 100, message=FALSE) -> tentilla_sim
print("SIMMAP Tentilla2")
make.simmap(ultratree, tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
print("SIMMAP Nectosome")
make.simmap(ultratree, nectosome, nsim = 100, message=FALSE) -> nectosome_sim
print("SIMMAP Nectosome Position")
make.simmap(ultratree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
print("SIMMAP Nectophore Types")
make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
print("SIMMAP Habitat")
make.simmap(ultratree, vrange, nsim = 100, message=FALSE) -> vrange_sim

@@

SIMMAP Sex distribution:
```{r sex, echo=F, message=F, warning=F}
#Simmap SEX
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Dioecious", "Monoecious")
sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(sex_sim)
```

Posterior probabilities of states (0 Dioecious, 1 Monoecious).

SIMMAP Presence of palpons:
```{r palpons, echo=F, message=F, warning=F}
#Simmap PALPONS
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(palpons_sim)
```

Posterior probabilities of states (1 Present, 0 Absent).

SIMMAP Presence of tentilla:
```{r tentilla, echo=F, message=F,warning=F}
#Simmap TENTILLA
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
tentilla_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(tentilla_sim)
```

Posterior probabilities of states (1 Present, 0 Absent).

@@

SIMMAP Presence of tentilla - *Physalia* corrected:
```{r tentilla2, echo=F, message=F, warning=F}
#Simmap TENTILLA
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
tentilla_sim2 %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(tentilla_sim2)
```

Posterior probabilities of states (1 Present, 0 Absent).

SIMMAP Presence of Nectophores:
```{r nectosome, echo=F, message=F, warning=F}
#Simmap NECTOSOME
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(nectosome_sim)
```

Posterior probabilities of states (1 Present, 0 Absent).

SIMMAP Position of Nectosome:
```{r nectosome_position, echo=F, message=F, warning=F}
#SIMMAP NECTOSOME POSITION
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green")
names(colors) = c("Dorsal", "No nectosome", "Ventral")
necto_pos_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

SIMMAP Nectophore number and types:
```{r nectophore_number_types, echo=F, message=F, warning=F}
#Simmap NECTOPHORE TYPES
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green", "blue")
names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
necto_types_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

Posterior probabilities of states (1 Present, 0 Absent).

It is not clear whether or not the stem group of siphonophores had a pneumatophore.

SIMMAP Habitat:
```{r depth_zones, echo=F, message=F, warning=F}
#Simmap Vertical.Range
plotTree(ultratree, lwd = 4)
colors = c("black", "cyan", "blue", "green", "red")
names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
vrange_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

Continuous traits: Brownian motion reconstructions

Median depth reconstruction:
```{r median_cont, echo=F, message=F, warning=F}
depth_value = depth_median
depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
plot(depthmap)
```

Bathymetrics phylomorphospace
```{r depth_phylomorpho, echo=F, message=F, warning=F}
par(oma=rep(3,4))
phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=1, label = "horizontal")
```

Bathymetrics phylomorphospace with time from root as color
```{r depth_phylochronomorpho, echo=F, message=F, warning=F}
evotime = fastBM(dpruned_tree,nsim=2)
evotime_map <- contMap(dpruned_tree,evotime[,1], plot=F)
evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=1, label = "horizontal")
#add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
```

Abundance of each taxon in the Monterey Bay sampling site
```{r counts_bars, echo=F, message=F, warning=F}
par(oma=rep(1,4))
log_abundance = log(abundance)
log_abundance["Physalia physalis"] = 0
plotBranchbyTrait(dpruned_tree, log_abundance, mode = "tips")
```

Uncertainty inclusive traitgram for depth range
```{r uncertainty_depth, echo=F, message=F, warning=F}
contTrait <- depth_range
A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)
paintree<-paintSubTree(dpruned_tree,node=length(dpruned_tree$tip)+1,"1")
trans<-as.character(floor(0:50/2))
trans[as.numeric(trans)<10]<- paste("0", trans[as.numeric(trans)<10],sep="")
# plot
for(i in 0:50){
  p<-i/length(trans)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,1]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=i>0)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,2]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=TRUE)
}
phenogram(dpruned_tree,c(contTrait,A$ace),add=TRUE, colors=setNames("black",1))
```

Bathymetrics scattergram array
```{r depth_scattergram, echo=F, message=F, warning=F}
fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
```

Distribution of sex across siphonophore taxa
```{r ggtree_sex, echo=F, message=F, warning=F}
exp_tree = ultratree
exp_tree$edge.length = 10*ultratree$edge.length
dd = data.frame(taxa  = tree$tip.label, trait = sex[match(tree$tip.label, names(sex))])
p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
p+theme(legend.position="right") + guides(color=guide_legend(title="Sex distribution", reverse=T)) 
```

Tentilla presence across siphonophore taxa
```{r ggtree_tentillq, echo=F, message=F, warning=F}
exp_tree = ultratree
exp_tree$edge.length = 10*ultratree$edge.length
dd = data.frame(taxa  = tree$tip.label, trait = as.factor(tentilla[match(tree$tip.label, names(tentilla))]))
p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
p+theme(legend.position="right") + guides(color=guide_legend(title="Tentilla", reverse=T)) 
```

Palpon presence across siphonophore taxa
```{r ggtree_palpon, echo=F, message=F, warning=F}
exp_tree = ultratree
exp_tree$edge.length = 10*ultratree$edge.length
dd = data.frame(taxa  = tree$tip.label, trait = as.factor(palpons[match(tree$tip.label, names(palpons))]))
p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
p+theme(legend.position="right") + guides(color=guide_legend(title="Palpons", reverse=T)) 
```

Overview grid of main binary traits
```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
p <- ggtree(ultratree)
p <- p + geom_tiplab(size=3)
f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:7)]
rownames(f_cdata) = rownames(cdata)
names(f_cdata)[4] = "Float"
gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 3)
```

Phylogenetic signal in the character data
Agnostic branch length tree generation:
```{r flat_tree}
flat_tree = tree
flat_tree$edge.length = rep(mean(flat_tree$edge.length), length(flat_tree$edge.length))
```

```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}
binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5,6,7)]), ultratree)
```
Phylogenetic signal in Binary Traits
```{r phylosig_binaries, echo=F, message=F, warning=F}
binaries_phylosig
```

```{r phylosig_depth_declare, message=F, warning=F, include=FALSE}
depth_physig = physignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
depth_multisig = multiPhylosignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
```
Phylogenetic signal in VARS bathymetrics
```{r phylosig_depth, echo=F, message=F, warning=F}
depth_physig
depth_multisig
```

Phylogenetic PCA for bathymetrics
```{r pPCA_depth, echo=F, message=F, warning=F}
phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") %>% biplot

@@

Phylogenetic ANOVA for depth
```{r phylANOVA_depth_declare, echo=F, message=F, warning=F, include=F}
z_nectopos = phylANOVA(dpruned_tree, dpruned_data$Nectosome.position, as.numeric(depth_median))
z_sex = phylANOVA(dpruned_tree, dpruned_data$Sex, as.numeric(depth_median))
z_necto = phylANOVA(dpruned_tree, dpruned_data$Nectosome, as.numeric(depth_median))
z_nectotype = phylANOVA(dpruned_tree, dpruned_data$Nectophores, as.numeric(depth_median))
z_palp = phylANOVA(dpruned_tree, dpruned_data$Palpons, as.numeric(depth_median))
z_tentilla = phylANOVA(dpruned_tree, dpruned_data$Tentilla, as.numeric(depth_median))
z_float = phylANOVA(dpruned_tree, dpruned_data$Pneumatophore, as.numeric(depth_median))
```
Each phylogenetic ANOVA compares variance of depth (continuous variable) within and between groups clustered by their categorical biological traits.
```{r phylANOVA_depth, echo=F, message=F, warning=F}
print("Depth vs Sex Distribution")
z_sex
print("Depth vs Nectosome")
z_necto
print("Depth vs Nectophore Types and Number")
z_nectotype
print("Depth vs Nectosome Position")
z_nectopos
print("Depth vs Palpons")
z_palp
print("Depth vs Tentilla")
z_tentilla
print("Depth vs Pneumatophore")
z_float
```

PGLS - Sex distribution
```{r pgls_depth_sex, echo=F, message=F, warning=F}
dp_sex <- dpruned_data$Sex
dp_sex = as.numeric(dp_sex)
names(dp_sex) = names(depth_median)
dp_sex[dp_sex==2] = 0        # 0 = Monoecious, 1 = Dioecious
#Vx <- Vy <- Cxy <- rep(0,length(dpruned_tree$tip.label))
#pgls.Ives(dpruned_tree, dp_sex, depth_median, Vx, Vy, Cxy)
pGLS = pgls.Ives(dpruned_tree, rep(dp_sex,3), rep(depth_median,3))
phylolm(depth_median~dp_sex,data=as.data.frame(cbind(depth_median,dp_sex)),phy=dpruned_tree,model="BM",measurement_error=TRUE,boot=100)
gls(dp_sex ~ depth_median)
```

>>>>

In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids. Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix and phylogeny due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes). ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution, and bootstrap values were estimated using 1000 replicates. BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112]. Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model. The CAT-GTR and CAT-poisson models did not converge after a long CPU time, and only the results from the CAT-poisson model are included here.

@@

#####The Tree

```{r ultratree, echo=F, message=F, warning=F}
par(ask=F)
par(mfrow=c(1,1))
plotTree(ultratree, lwd = 4)

```
```{r simmap_making, echo=F, message=F, warning=F}
#Stochastic trait mapping:
#Make simmap objects
#print("SIMMAP Sex Distribution")
make.simmap(ultratree, sex, nsim = 100, message=FALSE) -> sex_sim
#print("SIMMAP Palpons")
make.simmap(ultratree, palpons, nsim = 100, message=FALSE) -> palpons_sim
#print("SIMMAP Tentilla")
make.simmap(ultratree, tentilla, nsim = 100, message=FALSE) -> tentilla_sim
#print("SIMMAP Tentilla2")
make.simmap(ultratree, tentilla2, nsim = 100, message=FALSE) -> tentilla_sim2
#print("SIMMAP Nectosome")
make.simmap(ultratree, nectosome, nsim = 100, message=FALSE) -> nectosome_sim
#print("SIMMAP Nectosome Position")
make.simmap(ultratree, necto_pos, nsim = 100, message=FALSE) -> necto_pos_sim
#print("SIMMAP Nectophore Types")
make.simmap(ultratree, necto_types, nsim = 100, message=FALSE) -> necto_types_sim
#print("SIMMAP Habitat")
make.simmap(ultratree, vrange, nsim = 100, message=FALSE) -> vrange_sim

@@

#####SIMMAP Sex distribution
```{r sex, echo=F, message=F, warning=F}
#Simmap SEX
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Dioecious", "Monoecious")
sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(sex_sim)
```
Posterior probabilities of states (0 Dioecious, 1 Monoecious).

#####SIMMAP Presence of palpons
```{r palpons, echo=F, message=F, warning=F}
#Simmap PALPONS
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(palpons_sim)
```
Posterior probabilities of states (1 Present, 0 Absent).

#####SIMMAP Presence of tentilla
```{r tentilla, echo=F, message=F,warning=F}
#Simmap TENTILLA
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
tentilla_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(tentilla_sim)
```
Posterior probabilities of states (1 Present, 0 Absent).

@@

#####SIMMAP Presence of tentilla - *Physalia* corrected
```{r tentilla2, echo=F, message=F, warning=F}
#Simmap TENTILLA
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
tentilla_sim2 %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(tentilla_sim2)
```
Posterior probabilities of states (1 Present, 0 Absent).

#####SIMMAP Presence of Nectophores
```{r nectosome, echo=F, message=F, warning=F}
#Simmap NECTOSOME
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(nectosome_sim)
```
Posterior probabilities of states (1 Present, 0 Absent).

#####SIMMAP Position of Nectosome
```{r nectosome_position, echo=F, message=F, warning=F}
#SIMMAP NECTOSOME POSITION
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green")
names(colors) = c("Dorsal", "No nectosome", "Ventral")
necto_pos_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

#####SIMMAP Nectophore number and types
```{r nectophore_number_types, echo=F, message=F, warning=F}
#Simmap NECTOPHORE TYPES
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green", "blue")
names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
necto_types_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```
Posterior probabilities of states (1 Present, 0 Absent).It is not clear whether or not the stem group of siphonophores had a pneumatophore.

#####SIMMAP Habitat
```{r depth_zones, echo=F, message=F, warning=F}
#Simmap Vertical.Range
plotTree(ultratree, lwd = 4)
colors = c("black", "cyan", "blue", "green", "red")
names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
vrange_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

####Continuous traits: Brownian motion reconstructions

#####Median depth reconstruction:
```{r median_cont, echo=F, message=F, warning=F}
depth_value = depth_median
depthmap <- contMap(dpruned_tree, depth_value, plot=F) %>% setMap(colors=c("cyan","blue","black"))
plot(depthmap)
```

#####Bathymetrics phylomorphospace
```{r depth_phylomorpho, echo=F, message=F, warning=F}
par(oma=rep(3,4))
phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=1, label = "horizontal")
```

#####Bathymetrics phylomorphospace with time from root as color
```{r depth_phylochronomorpho, echo=F, message=F, warning=F}
evotime = fastBM(dpruned_tree,nsim=2)
evotime_map <- contMap(dpruned_tree,evotime[,1], plot=F)
evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=1, label = "horizontal")
#add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
```

####Abundance of each taxon in the Monterey Bay sampling site
```{r counts_bars, echo=F, message=F, warning=F}
par(oma=rep(1,4))
log_abundance = log(abundance)
log_abundance["Physalia physalis"] = 0
plotBranchbyTrait(dpruned_tree, log_abundance, mode = "tips")
```

###Uncertainty inclusive traitgram for depth range
```{r uncertainty_depth, echo=F, message=F, warning=F}
contTrait <- depth_range
A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)
paintree<-paintSubTree(dpruned_tree,node=length(dpruned_tree$tip)+1,"1")
trans<-as.character(floor(0:50/2))
trans[as.numeric(trans)<10]<- paste("0", trans[as.numeric(trans)<10],sep="")
# plot
for(i in 0:50){
  p<-i/length(trans)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,1]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=i>0)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,2]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=TRUE)
}
phenogram(dpruned_tree,c(contTrait,A$ace),add=TRUE, colors=setNames("black",1))
```

####Bathymetrics scattergram array
```{r depth_scattergram, echo=F, message=F, warning=F}
fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
```

####Distribution of sex across siphonophore taxa
```{r ggtree_sex, echo=F, message=F, warning=F}
exp_tree = ultratree
exp_tree$edge.length = 10*ultratree$edge.length
dd = data.frame(taxa  = tree$tip.label, trait = sex[match(tree$tip.label, names(sex))])
p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
p+theme(legend.position="right") + guides(color=guide_legend(title="Sex distribution", reverse=T)) 
```

####Tentilla presence across siphonophore taxa
```{r ggtree_tentillq, echo=F, message=F, warning=F}
exp_tree = ultratree
exp_tree$edge.length = 10*ultratree$edge.length
dd = data.frame(taxa  = tree$tip.label, trait = as.factor(tentilla[match(tree$tip.label, names(tentilla))]))
p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
p+theme(legend.position="right") + guides(color=guide_legend(title="Tentilla", reverse=T)) 
```

####Palpon presence across siphonophore taxa
```{r ggtree_palpon, echo=F, message=F, warning=F}
exp_tree = ultratree
exp_tree$edge.length = 10*ultratree$edge.length
dd = data.frame(taxa  = tree$tip.label, trait = as.factor(palpons[match(tree$tip.label, names(palpons))]))
p = ggtree(exp_tree, layout="rectangular") + xlim(NA, 23)
p = p %<+% dd + geom_tiplab(aes(color=trait), show_guide = FALSE, size = 3.5) + geom_tippoint(aes(color=trait), alpha=0.25)
p+theme(legend.position="right") + guides(color=guide_legend(title="Palpons", reverse=T)) 
```

#####Overview grid of main binary traits
```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
p <- ggtree(ultratree)
p <- p + geom_tiplab(size=3)
f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:7)]
rownames(f_cdata) = rownames(cdata)
names(f_cdata)[4] = "Float"
gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 3)
```

####Phylogenetic signal in the character data

#####Agnostic branch length tree generation:
```{r flat_tree}
flat_tree = tree
flat_tree$edge.length = rep(mean(flat_tree$edge.length), length(flat_tree$edge.length))
```

```{r phylosig_binaries_declare, echo=F, message=F, warning=F, include=F}
binaries_phylosig = multiPhylosignal(as.matrix(cdata[,c(2,5,6,7)]), ultratree)
```
#####Phylogenetic signal in Binary Traits
```{r phylosig_binaries, echo=F, message=F, warning=F}
binaries_phylosig
```

```{r phylosig_depth_declare, message=F, warning=F, include=FALSE}
depth_physig = physignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
depth_multisig = multiPhylosignal(dpruned_data[,c(14,15,16,18,19)], dpruned_tree)
```
#####Phylogenetic signal in VARS bathymetrics
```{r phylosig_depth, echo=F, message=F, warning=F}
depth_physig
depth_multisig
```

#####Phylogenetic PCA for bathymetrics
```{r pPCA_depth, echo=F, message=F, warning=F}
phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") %>% biplot

@@

#####Phylogenetic ANOVA for depth
```{r phylANOVA_depth_declare, echo=F, message=F, warning=F, include=F}
z_nectopos = phylANOVA(dpruned_tree, dpruned_data$Nectosome.position, as.numeric(depth_median))
z_sex = phylANOVA(dpruned_tree, dpruned_data$Sex, as.numeric(depth_median))
z_necto = phylANOVA(dpruned_tree, dpruned_data$Nectosome, as.numeric(depth_median))
z_nectotype = phylANOVA(dpruned_tree, dpruned_data$Nectophores, as.numeric(depth_median))
z_palp = phylANOVA(dpruned_tree, dpruned_data$Palpons, as.numeric(depth_median))
z_tentilla = phylANOVA(dpruned_tree, dpruned_data$Tentilla, as.numeric(depth_median))
z_float = phylANOVA(dpruned_tree, dpruned_data$Pneumatophore, as.numeric(depth_median))
```
Each phylogenetic ANOVA compares variance of depth (continuous variable) within and between groups clustered by their categorical biological traits.

```{r phylANOVA_depth, echo=F, message=F, warning=F}
print("Depth vs Sex Distribution")
z_sex
print("Depth vs Nectosome")
z_necto
print("Depth vs Nectophore Types and Number")
z_nectotype
print("Depth vs Nectosome Position")
z_nectopos
print("Depth vs Palpons")
z_palp
print("Depth vs Tentilla")
z_tentilla
print("Depth vs Pneumatophore")
z_float
```

#####PGLS - Sex distribution
```{r pgls_depth_sex, echo=F, message=F, warning=F}
dp_sex <- dpruned_data$Sex
dp_sex = as.numeric(dp_sex)
names(dp_sex) = names(depth_median)
dp_sex[dp_sex==2] = 0        # 0 = Monoecious, 1 = Dioecious
#Vx <- Vy <- Cxy <- rep(0,length(dpruned_tree$tip.label))
#pgls.Ives(dpruned_tree, dp_sex, depth_median, Vx, Vy, Cxy)
pGLS = pgls.Ives(dpruned_tree, rep(dp_sex,3), rep(depth_median,3))
phylolm(depth_median~dp_sex,data=as.data.frame(cbind(depth_median,dp_sex)),phy=dpruned_tree,model="BM",measurement_error=TRUE,boot=100)
gls(dp_sex ~ depth_median)
```
